How to obtain the redshift distribution from probabilistic redshift estimates

### 2017-06-26 research/paper/draft.tex
merged test cases
math added

Before jumping into the details of the model, it is crucial to settle on an 
interpretation of what a photo-$z$ PDF actually is.  When a photo-$z$ PDF is 
reported, it is an object containing information about galaxy $i$'s redshift 
$z_{i}$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF a 
\textit{posterior} probability distribution.  The data $\vec{d}_{i}$ may be 
fluxes, magnitudes, colors, or any combination thereof.  We assume that the 
redshifts of the galaxies are not related to the redshifts or photometry of 
other galaxies in the survey, i.e. they are \textit{independent}.

@@

According to Bayes' rule, the posterior distribution of interest is
\begin{align}
\log p(\vec{\phi} | \{\vec{d}_{i}\}) &\propto \log p(\{\vec{d}_{i}\} | 
\vec{\phi}) + \log p(\vec{\phi}),
\end{align}
where $p(\vec{\phi})$ is the \textit{hyperprior} probability distribution over 
possible values of the hyperparameters.  The hyperprior is a choice made by 
those performing the inference and is not to be confused with the interim prior 
parameters that influence the photo-$z$ PDF data product.

Focusing on the likelihood term $p(\{\vec{d}_{i}\} | \vec{\phi})$, we employ 
the power of the hierarchical model through \textit{marginalization} of the 
unobserved redshift parameters by integrating over them to obtain 
\begin{align}
\log[p(\{\vec{d}_{i}\} | \vec{\phi})] &= \log[\int\exp[\log[p(\{\vec{d}_{i}\} | 
\{z_{i}\})]\\
& + \log[p(\{z_{i}\} | \vec{\phi})]]\ d\{z_{i}\}].
\end{align}

Given a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, 
\vec{\alpha})\}$

To make use of the photo-$z$ interim priors, it is absolutely essential that 
$\vec{\alpha}$ be known.

This framework entails a number of choices and assumptions that must be 
addressed explicitly.

\begin{enumerate}
	\item The chosen functional form with parameters $\vec{\phi}$ must 
actually be able to describe the true redshift distribution.
	\item The photometric data and redshift for each galaxy is independent 
of the photometric data and redshifts of all other galaxies.
	\item There is one interim prior with parameters $\vec{\phi}^{*}$ 
shared among all galaxies in the survey.
	\item We must choose a hyperprior probability distribution 
$p(\vec{\phi})$ over the hyperparameters.
	\item While we advocate for the approach of hierarchical inference, the 
probabilistic graphical model presented here is not the only one that could be 
proposed.
\end{enumerate}

@@

Because the realistic mock data of Sec. \ref{sec:buzzard} is associated with 
true values of the cosmological parameters, we may compare the quality of 
cosmological constraints under different estimators of $n(z)$, creating a 
figure like Fig. \ref{eq:scgsr}.  We perform this analysis using a forecasting 
code that takes $n(z)$ and produces projected error ellipses in the space of 
cosmological parameters.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{cosmo_constraints.png}
		\caption{[I'd like to be able to make a plot like this with the 
different $n(z)$ estimators produced by \texttt{chippr} from the same set of 
photo-$z$ PDFs.  Then I can calculate some metrics of the accuracy and 
precision of the error distributions relative to the true values of the 
cosmological parameters that produced the mock data.]}
		\label{fig:scgsr}
	\end{center}
\end{figure}

@@

Thee results of Fig \ref{fig:money} have significant implications for the 
developing data analysis pipelines of next-generation telescope surveys.  
However, the method presented here has its own limitations, which are 
reiterated to discourage the community from applying this work inappropriately.

>>>>

Before jumping into the details of the model, it is crucial to settle on an 
interpretation of what a photo-$z$ PDF actually is.  When a photo-$z$ PDF is 
reported, it is an object containing information about the redshift $z_{i}$ of 
galaxy $i$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF 
a \textit{posterior} probability distribution.  The data $\vec{d}_{i}$ may be 
fluxes, magnitudes, colors, or any combination thereof.  We assume that the 
redshifts of the galaxies are not related to the redshifts or photometry of 
other galaxies in the survey, i.e. they are \textit{independent}.

@@

According to Bayes' rule, the posterior distribution of interest is
\begin{align}
\label{eq:hyper_bayes}
\log[p(\vec{\phi} | \{\vec{d}_{i}\})] &\propto \log[p(\{\vec{d}_{i}\} | 
\vec{\phi})] + \log[p(\vec{\phi})] ,
\end{align}
where $p(\vec{\phi})$ is the \textit{hyperprior} probability distribution over 
possible values of the hyperparameters.  The hyperprior is a choice made by 
those performing the inference and is not to be confused with the interim prior 
parameters that influence the photo-$z$ PDF data product.

Focusing on the likelihood term $p(\{\vec{d}_{i}\} | \vec{\phi})$, we employ 
the power of the hierarchical model through \textit{marginalization} of the 
unobserved redshift parameters by integrating over them to obtain 
\begin{align}
\label{eq:marginalize}
\begin{split}
\log[p(\{\vec{d}_{i}\} | \vec{\phi})] = & 
\log\left[\int\exp\left[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\right.\right. \\
& \left.\left.+ \log[p(\{z_{i}\} | \vec{\phi})]\right]\ d\{z_{i}\}\right] .
\end{split}
\end{align}
The redshifts $\{z_{i}\}$ are statistically independent of each other, and each 
galaxy's photometry depends solely on its own redshift, so we may invoke 
statistical independence to get
\begin{align}
\label{eq:independence_hyper}
\log[p(\{z_{i}\} | \vec{\phi})] &= \sum_{i}^{N} \log[p(z_{i} | \vec{\phi})]
\end{align}
and
\begin{align}
\label{eq:independence}
\log[p(\{\vec{d}_{i}\} | \{z_{i}\})] &= \sum_{i}^{N} \log[p(\vec{d}_{i} | 
z_{i})] .
\end{align}
Having chosen the functional form mapping $\vec{\phi}$ to $n(z)$, the terms on 
the righthand side of Eq. \ref{eq:independence_hyper} are known.  However, the 
terms on the righthand side of Eq. \ref{eq:independence} are likelihoods, and 
the photo-$z$ PDFs we have are interim posteriors.

To transform the interim posteriors we have into the likelihoods we need, we 
start with the vacuously true statement
\begin{align}
\begin{split}
\label{eq:unity}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})].
\end{split}
\end{align}
We expand the last term using Bayes' Rule to get
\begin{align}
\begin{split}
\label{eq:bayes}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{\phi}^{*})] - \log[p(\vec{d}_{i} | z_{i}, 
\vec{\phi}^{*})] + \log[p(\vec{d}_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Fig. \ref{fig:pgm} has no direct connection between $\vec{d}_{i}$ and 
$\vec{\phi}$, so we can break up $\log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]$ 
as follows:
\begin{align}
\begin{split}
\label{eq:split}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{\phi}^{*})] - \log[p(\vec{d}_{i} | z_{i})] - 
\log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Canceling terms, we find
\begin{align}
\label{eq:cancel}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] - 
\log[p(z_{i} | \vec{\phi}^{*})].
\end{align}

Now we combine Eqs. \ref{eq:hyper_bayes}, \ref{eq:marginalize}, 
\ref{eq:independence_hyper}, \ref{eq:independence}, and \ref{eq:cancel} to 
arrive at 
\begin{align}
\begin{split}
\label{eq:final}
\log[p(\vec{\phi} | \{\vec{d}_{i}\})] \propto& \log[p(\vec{\phi})] + 
\log\left[\int \exp\left[\sum_{i}^{N} \left(\right.\right.\right.\\
& \left.\left.\left. + \log[p(z_{i} | \vec{\phi})] - \log[p(z_{i} | 
\vec{\phi}^{*})]\right.\right.\right.\\
& \left.\left.\left. + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]
\right)\right] d\{z_{i}\}\right].
\end{split}
\end{align}
Thus, if we have a catalog of photo-$z$ interim posteriors $\{p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})\}$, the interim prior parameters $\vec{\phi}^{*}$, 
and a prior distribution $p(\vec{\phi})$, we may find the posterior 
$\log[p(\vec{\phi} | \{\vec{d}_{i}\})]$ on the redshift distribution $n(z)$.

This framework entails a number of choices and assumptions that must be 
addressed explicitly:
\begin{enumerate}
	\item The chosen functional form with parameters $\vec{\phi}$ must be 
capable of describing the true redshift distribution.
	\item The photometric data and redshift for each galaxy is independent 
of the photometric data and redshifts of all other galaxies.
	\item There is one interim prior shared among all galaxies in the 
survey, and its parameters with known parameters $\vec{\phi}^{*}$ are known.
	\item We must choose a hyperprior probability distribution 
$p(\vec{\phi})$ over the hyperparameters.
	\item While we advocate for the approach of hierarchical inference, the 
probabilistic graphical model presented here is not the only one that could be 
proposed.
\end{enumerate}

@@

Because the realistic mock data of Sec. \ref{sec:buzzard} is associated with 
true values of the cosmological parameters, we may compare the quality of 
cosmological constraints under different estimators of $n(z)$, creating a 
figure like Fig. \ref{fig:scgsr}.  We perform this analysis using a forecasting 
code that takes $n(z)$ and produces projected error ellipses in the space of 
cosmological parameters.

\begin{figure}
	\begin{center}
		Not my figure! (N. MacCrann via J. DeRose)
		\includegraphics[width=0.5\textwidth]{cosmo_constraints.png}
		\caption{[I'd like to be able to make a plot like this with the 
different $n(z)$ estimators produced by \texttt{chippr} from the same set of 
photo-$z$ PDFs.  Then I can calculate some metrics of the accuracy and 
precision of the error distributions relative to the true values of the 
cosmological parameters that produced the mock data.]}
		\label{fig:scgsr}
	\end{center}
\end{figure}

@@

The results of Fig \ref{fig:money} have significant implications for the 
developing data analysis pipelines of next-generation telescope surveys.  
However, the method presented here has its own limitations, which are 
reiterated to discourage the community from applying this work inappropriately.
